---
title: "Processing of COMPASS-FME Data"
author: "A Hart"
date: "2024-07-23"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background
### COMPASS-FME Overview
COMPASS-FME is a project studying how coastal ecosystems respond to stress from rising sea levels and storm surges.  The project collects data from seven different sites in the Chesapeake Bay and Western Lake Erie Basin.  __The sites all have differing salinity levels and soil types to look a wide range of ecosystems.__  Each site has three plots: upland, transition zone, and wetland area.

<img src ="C:/Users/hart187/OneDrive - PNNL/Documents/compass-sites.png">

In addition to multiple sites, COMPASS-FME uses multiple data sensors.  The sensors fall into one of the following: Weather(rain, temperature, etc), Soil (soil electrical conductivity, soil temperature, etc), Open Soil(water pressure, water salinity, etc), Ground Water(groundwater salinity, groundwater density, etc), or Vegetation(sapflow, etc).

- *(What COMPASS is, Sites and Sensors)*

### Data
COMPASS-FME generates over 3 million data per month.  That is a LOT of data.  The data is stored in folders separated by site and year.  Within these folders, there are several csv files holding the data.  The v1 version, the version the processing process is tailored too, has 12 columns.  The columns needed for processing are **Site, Plot, TIMESTAMP, Value, and research_name**.

```{r v1 data, echo = FALSE, message=FALSE}
library(readr)
b <- read_csv("C:/Users/hart187/COMPASS_folders/CRC_2022/CRC_TR_20220401-20220430_L1_v1-0.csv", show_col_types = FALSE)

print(head(b))
```


- *(How data is Stored; LOTS of data)*

## Processing
### HPC
- Explain how the HPC works

### Summarizing script
- talk about script that summarizes data

## Post-Processing {.tabset}
- explain purpose of post processing script
- different graph types

### Average
- Explain setup of graphs (split by plot and geom_smooth)

### Median Stardard Deviation
- Explain what mad is
- explain setup and outliers function

### Percentage of NA
- explain the reason for computing this
- setup

### Standard Deviation
- explain setup
